---
title: "Covid91 vaccine study Final2022"
author: "Cesar Guillen"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(FinalDat, sex=="M")
```

```{r}
Males <- select(Males, -sex)
```

## Barplot for Males
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

As seen in both graphs we can infer that the vaccine is working better than the placebo for the Males in the expiriment because there have been fewer cases where those who were given the vaccine contracted COVID19 than those who took the placebo. 

## Numerical Summary Male
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
This chart is showing that the percent of males who took the placebo and were infected with COVID19 was 2.68 percent, while the percent of males who took the vaccine and were infected with COVID19 was 1.28 percent, which we can guess means that the vaccine is working efficiently for males. 

## Inferential Results Male
```{r}
chisq.test(table1)
chisqtestGC(table1)
fishtestm<-fisher.test(table1)
print(fishtestm)
```

According to the inferential results above, we can see that the men who took the placebo are almost twice as likely to contract COVID19 than the men who took the vaccine. And with the p-value being 2.0921 we can assume that the chances are low for there being no difference between the placebo and the vaccine.

## Efficancy Males
```{r}
options(digits = 1)
maleffic=fishtestm$estimate/(1+fishtestm$estimate)*100
malefficlo=fishtestm$conf.int[1]/(1+fishtestm$conf.int[1])*100
maleffichi=fishtestm$conf.int[2]/(1+fishtestm$conf.int[2])*100
print(maleffic)
print(malefficlo)
print(maleffichi)
```
The confidence interval for the effectiveness for the vaccine for the males in the expirement has a low of `r malefficlo`, and a high of `r maleffichi`. 




# Females

```{r}
Females <- filter(FinalDat, sex=="F")
```

```{r}
Females <- select(Females, -sex)
```

## Barplot for Females
```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Both box plots above show that there have been very few cases where the females who have been given the placebo and the vaccine where they have gotten COVID19, but it shows that those who took the vaccine are less likley to get COVID19 than those who took the placebo. 

## Numerical Summary Females
```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```
According to the table, 2.3 percent of the females who took the placebo were infected with COVID19, which is almost twice the amount of those who got COVID19 and were given the vaccine, which can be useful information as to determine which is better for females. 

## Inferential Summary Females
```{r}
chisq.test(table2)
chisqtestGC(table2)
fishtestf<-fisher.test(table2)
fishtestf
```
The table above gives us two very promising signs of the vaccine, one that shows the ratio of those who took the placebo are 1.42 percent more likely to contract the virus than those who took the vaccine, which would mean that the vaccine would work for females as well. We are also given a small p-value, which is .0009484, meaning the chances of getting no difference between the vaccine and the placebo are extremely small. 

## Efficacy Females
```{r}
options(digits = 1)
femeffic=fishtestf$estimate/(1+fishtestf$estimate)*100
femefficlo=fishtestf$conf.int[1]/(1+fishtestf$conf.int[1])*100
femeffichi=fishtestf$conf.int[2]/(1+fishtestf$conf.int[2])*100
print(femeffic)
print(femefficlo)
print(femeffichi)
```
The confidence interval of the effectiveness of the vaccine for females has a low `r femefficlo`, and has a high of `r femeffichi`. 


# LGBTQ



# Druggies

